Funding programmes for early career researchers do not improve their independence

### 2024-01-19 article_formatted.Rmd

---
title: "Mission Unaccomplished: A Funding Programme for Early Career Researchers Fails to Promote Independence"
author:
- name: David Janků
  affiliation: '[1] Centre for Science, Technology, and Society
    Studies, Institute of Philosophy, Czech Academy of Sciences; [2] Institute of Sociological Studies, Faculty of Social Sciences,
    Charles University'
- name: Radim Hladík
  affiliation: '[1] Centre for Science, Technology, and Society Studies, Institute
    of Philosophy, Czech Academy of Sciences'
output: bookdown::html_document2
  # pdf_document: default
  # html_document:

@@

This work was supported by the Czech Science Foundation (project no. GJ20-01752Y, _Funded and Unfunded Research in the Czech Republic: Scientometric Analysis and Topic Modeling_).

@@

Name: David Janků 
Affiliation: 
[1] Centre for Science, Technology, and Society Studies, Institute of Philosophy, Czech Academy of Sciences
[2] Institute of Sociological Studies, Faculty of Social Sciences, Charles University
Address: Jilská 1, 110 00, Praha 1, Czech Republic  
E-mail: david.janku@fsv.cuni.cz

@@

Independence is an important factor in research excellence, often assessed for hiring, tenure and promotion in academia.
From a macro perspective, increased independence might also lead to faster exploration of the cognitive map of science and thus accelerate innovation.
Many national research funders have created programmes specifically aimed at increasing independence of early career researchers.
A quantifiable measure of such programmes' impact on changes in independence has long been lacking..
We employed a recently proposed independence indicator [@vandenbesselaarMeasuringResearcherIndependence2019] to asses whether one such funding programme operated by Czech Science Foundation succeeds at stimulating independence.
We found that the e funding scheme designed for early career researchers is not effective at stimulating their independence.
Regardless of the type of funding, the strongest effect on independence growth is time, explaining around 60% of the variance in independence.
Additionally, we found differences in levels of independence across disciplines that were stable across time and treatment groups, with Social Sciences and Humanities showing highest independence, while  Agricultural and Veterinary sciences and Medical and Health Sciences showing lowest independence.
However, these differences reflect only changes in collaboration patterns, while there were no changes in cognitive patterns across disciplines and only modest changes across time (from 40 % to 46 % of topics not shared by their PhD advisor).

## Keywords
research independence; early career researchers; project funding; mentoring; counter-factual analysis; research evaluation

# Intro

## Intro 1: focus on the bigger picture, sociology and policy perspective

Researchers’ independence from their PhD advisor is one of the criteria often used for assessing researchers for hiring, tenure and promotion (@vanarensbergenDifferentViewsScholarly2014; @moherAssessingScientistsHiring2018), constituting one factor of research excellence.
The promise of independence is that researchers will “challenge the existing common views and open up new lines of research” (@vandenbesselaarMeasuringResearcherIndependence2019, p.2).

Early career researcher independence is central to the long-lasting debate about how science progresses - whether it is indeed guided mostly by the truth and empirical evidence, or whether more social and demographic factors might play an important role. @hull1978 found that even more senior researchers readily updated their beliefs with the new evidence, supporting the former, though many (famously including Max Planck) observed that senior researchers were more fixed in their beliefs and new ideas were mostly brought up by the incoming early career researchers giving rise to the “science moves forward one funeral at the time” proverb. More recently, @azoulay2010 provided further support for the latter by investigating the effects of unexpected deaths of prominent scientists on their research fields and their collaborators' careers. They found a decrease in the productivity of junior collaborators and an increase in published work by young newcomers to the field rewarded by an increased citation rate. That suggests the sudden deaths of prominent scientists gave room for more disruptive changes in the field and for bringing new ideas and perspectives while harming the status quo holders who worked with the prominent scientists on their research agendas - a “Goliath effect”, as coined by @wang2021, p.42).

/// here I can potentially talk a bit more about Inheritance of research agendas - how often does that happen, etc..///

Insofar as the latter is true, in order to accelerate scientific progress, we might want to tackle this trend by supporting early career researchers to gain more independence from their PhD advisors. Younger researchers are already more likely to come up with new topics and ideas, as found by @packalen2019 by looking at the usage of new words in research papers. However, they might still feel various kinds of pressures to stick with their mentor’s research agendas and lack of opportunities to become more independent. In recent decades, there has been a worry that researchers’ independence is decreasing, illustrated by an increase in age at which US biomedical researchers received their first grant (NAS, 2005; @danielsGenerationRiskYoung2015), as well as direct measures of new researchers becoming more dependent on their PhD advisors across almost all disciplines in Slovenia (@rojkoScientificPerformanceResearch2022, table 6 on p.9).

Perhaps to tackle this trend, many grant agencies decided to start awarding grants for early career researchers, usually up to 8 years after finishing their PhD. Such programmes have often explicitly stated goals of helping researchers become more independent, set up their own research group and open up new lines of research. The examples of such programmes could be Starting Grants by European Research Council (@StartingGrantERC), Junior project and Junior Star by Czech Science Foundation (2020), Walter N. Benjamin Programme by DFG (@WalterBenjaminProgramme), Innovation Research Incentives Scheme by Netherlands Organization of Scientific Research (NWO) (2021).

Moreover, stimulating independence and supporting researchers to open up new lines of research is also a worthy policy goal regardless of researchers’ age. @rzhetsky2015 found that biomedical researchers are in practice often choosing less exploratory strategies for topic choice than would be socially optimal if we wanted to explore the cognitive map of science (or in this case chemicals) as fast as possible, and their choices become more conservative over time. They suggest that increased interdisciplinarity, risk-taking and the publication of experimental failures would substantially improve the speed of discovery.

//// factors influencing independence////
There are few known factors influencing the level of research independence. One of them is study discipline. Based on some early empirical studies (@rojkoScientificPerformanceResearch2022), it seems that researchers in disciplines such as philosophy, psychology; religion, theology; social sciences; the arts, recreation, entertainment, sport; language, linguistics, literature; and geography, biography, history experience significantly higher levels of independence on their PhD advisor than researchers in disciplines such as science and knowledge, organization, computer science, information, documentation, librarianship, institutions, publications; mathematics, natural sciences; and applied sciences, medicine, technology.

Interestingly, there is also some evidence that the propensity to cognitive independence might be inherited and passed down from generation to generation in mentor-mentee relationships @yoshioka-kobayashi2021 - perhaps it is not only research agendas that students inherit from their advisors but also the tendency to deviate from those research agendas. This might also suggest that to increase junior researchers’ independence, we might want to teach their advisors to be more independent, in order to have a ripple effect down the line.

Finally, there is also some preliminary evidence that higher independence modestly correlates with the number of citations acquired (@rojkoScientificPerformanceResearch2022), but others argue that even the research systems with very low independence can create a lot of scientific impact, and thus the relationships between independence and impact might not be as clear (@glaserIndependenceResearchReview2022)

Although researcher independence seems to be an important feature that is acknowledged by many actors in the research systems and that research funders try to stimulate, there hasn’t been much effort focused on quantifying this phenomenon. Some studies try to measure independence in a simple way, as a ratio of publications the given researcher produced without their supervisor divided by the total number of publications of the given researchers (e.g. @rojkoScientificPerformanceResearch2022). However, such an indicator could be misleading in the world of team science, and also only focuses on collaborations, not tracking whether a given researcher becomes independent in their thinking and research topics. [@vandenbesselaarMeasuringResearcherIndependence2019] therefore proposed a more complex indicator for measuring researcher independence on the individual level, based on co-authorship network and topical overlaps. Their measure uses only information about publications’ characteristics (authorship and reference list) regardless of their impacts, unlike other alternative measures that focus solely on the impact of the independent work in terms of received citations (Dey, et al., 2021, @ALISNovelMetric).

Since the explicit goal of many funding instruments that are being used by research funders all over the world is to help early career researchers become independent, we decided to use the newly created indicators to assess the extent, to which such programmes influence researchers' independence.

## Intro 2: Perspective of “What are the effects of competitive grant funding? And looking at independence as one of such effects”

///Primer of grant-making context:

Grant funding mechanisms have previously been shown to stimulate productivity in terms of the number of publications, impact in terms of citations, and even scientific novelty (@wang2018, or as some authors argue rather interdisciplinarity: @fontana2020). In terms of researchers’ careers, obtaining grant funding was connected to significantly higher chances of career progression and obtaining a professorship (@bloch2014).

///Though there is still some disagreement about the direction and magnitude of some of these effects, c.f. @@glaserIndependenceResearchReview2022 on novelty. --> here I can potentially go more in-depth into the critiques of grant funding mechanisms? E.g. spell out the critiques, such that they are strengthening the Matthews effect, they reduce novelty, etc....

However, there are still several components of research excellence (see @vanarensbergenDifferentViewsScholarly2014; @moherAssessingScientistsHiring2018) that have not been investigated yet, with one having specifically important effects on researchers’ carers and development: their independence. Researchers’ independence from their PhD advisor is one of the criteria often used for assessing researchers for hiring, tenure and promotion (@vanarensbergenDifferentViewsScholarly2014; @moherAssessingScientistsHiring2018). The promise of independence is that researchers will “challenge the existing common views and open up new lines of research” (@vandenbesselaarMeasuringResearcherIndependence2019, p.2).

An independent investigator is then defined by the US National Academies of Science (NAS) as “one who enjoys independence of thought—the freedom to define the problem of interest and to choose or develop the best strategies and approaches to address that problem–especially during the earlier career phase” (@vandenbesselaarMeasuringResearcherIndependence2019, p.4). Such a definition does not necessarily put independence in relation to other actors (e.g. one’s PhD advisor), but only in relation to one’s own freedom and resources. However other institutions, such as European Research Council via their Starting Grants scheme (@neufeldPeerReviewbasedSelection2013) think of independence in a relational manner, in terms of having one (or a few) publications without the former PhD supervisor being co-author; being PI on a grant; or being lab director or research leader.

Based on some early empirical studies @rojkoScientificPerformanceResearch2022, it seems that researchers in disciplines such as philosophy, psychology; religion, theology; social sciences; the arts, recreation, entertainment, sport; language, linguistics, literature; and geography, biography, history experience significantly higher levels of independence on their PhD advisor than researchers in disciplines such as science and knowledge, organization, computer science, information, documentation, librarianship, institutions, publications; mathematics, natural sciences; and applied sciences, medicine, technology. 
There is also some preliminary evidence that higher independence modestly correlates with number of citations acquired (@rojkoScientificPerformanceResearch2022), but others argue that even the research systems with very low independence can create a lot of scientific impact, and thus the relationships between independence and impact might not be as clear (@glaserIndependenceResearchReview2022)

In recent decades, there has been a worry that researchers’ indepedence is decreasing, illustrated by increase in age at which US (biomedical) researchers received their first grant (NAS, 2005; @danielsGenerationRiskYoung2015), as well as direct measures of new researchers becoming more dependent on their PhD advisors across almost all disciplines in Slovenia (@rojkoScientificPerformanceResearch2022, table 6 on p.9). Perhaps to tackle this trend, many grant agencies decided to start awarding grants for early career researchers, usually up to 8 years after finishing their PhD. Such programmes have often explicitly stated goals of helping researchers become more independent, set up their own research group and open up new lines of research. The examples of such programmes could be Starting Grants by European Research Council (@StartingGrantERC), Junior project and Junior Star by Czech Science Foundation (2020), Walter N. Benjamin Programme by DFG (@WalterBenjaminProgramme), Innovation Research Incentives Scheme by Netherlands Organization of Scientific Research (NWO) (2021)... something from the US would be also good)

Although researcher independence seems to be important feature that is acknowledged by many actors in the research systems and that research funders try to stimulate, there hasn’t been much effort focused on quantifying this phenomenon. Some studies try to measure independence in a simple way, as a ratio of publications the given researcher produced without their supervisor divided by the total number of publications of the given researchers (e.g. @rojkoScientificPerformanceResearch2022). However, such indicator could be misleading in the world of team science, and also only focuses on collaborations, not tracking whether given researcher become independent in their thinking and research topics. [@vandenbesselaarMeasuringResearcherIndependence2019] therefore proposed more complex indicator for measuring researcher independence on the individual level, based on co-authorship network and topical overlaps. Their measure uses only information about publications’ characteristics (authorship and reference list) regardless of it’s impacts, unlike other alternative measures that focus solely on the impact of the independent work in terms of received citations (Dey, et al., 2021, @ALISNovelMetric).

Since the explicit goal of many funding instruments that are being used by research funders all over the world is to help early career researchers become independent, we used the newly created indicators to assess the extent, to which such programmes influence researchers independence.

H1: Researchers, who received funding from programmes aimed at funding early career researchers, will show more independence from their PhD advisors after receiving the funding than researchers who did not receive funding.

H2: Researchers, who received funding from programmes aimed at funding early career researchers, will show more independence from their PhD advisors compared to similar researchers who received other types of funding at similar point in theri career.

## Intro 3: focusing on support for early career researchers (and why is that important topic) and then mention the independence as one of the ways how we can test whether our support mechanisms for early stage researchers actually work

Funding for early career researchers is an important topic in science funding policy discussions as it might have a significant impact on career progression of the new generations of researchers (@wright2017; @pickett2019; @vandenbesselaar2015). At the same time, the age of receiving a first grant has been growing significantly over the past decades (@powell2016, @christian2021). These trends have motivated some to create new proposals for funding agencies and other research funders in order to support early career researchers (@dewinde2021). Some funding agencies decided to tackle this trend by giving advantage to first-awardees and early career researchers in their usual distribution process, while others have created specific programmes designated to funding young researchers soon after their PhD.

///maybe talk about other types of support that early career researchers get? Or maybe have another paragraph describing what does the life of early career researchers usually look like? (--> perhaps mentioning that the time to finish PhD gets longer and longer, people are now expected to do multiple post-docs, etc... --> perhaps borrow some info from Science of Science book Careers section)///

It is expected, that by giving early-stage researchers funding, they will be able to develop more independence and open up new lines of research (@vandenbesselaarMeasuringResearcherIndependence2019, p.2). Younger researchers are already more likely to come up with new topics and ideas, as found by [@packalen2019] by looking at the usage of new words in research papers. However, they might still feel various kinds of pressures to stick with their mentor’s research agendas and lack of opportunities to become more independent. In recent decades, there has been a worry that researchers’ independence is decreasing, illustrated by increase in age at which US (biomedical) researchers received their first grant (NAS, 2005; @danielsGenerationRiskYoung2015), as well as direct measures of new researchers becoming more dependent on their PhD advisors across almost all disciplines in Slovenia (@rojkoScientificPerformanceResearch2022, table 6 on p.9).
///talk about the balance between experience and independence??///

Perhaps to tackle this trend, many grant agencies decided to start awarding grants for early career researchers, usually up to 8 years after finishing their PhD. Such programmes have often explicitly stated goals of helping researchers become more independent, set up their own research group and open up new lines of research. The examples of such programmes could be Starting Grants by European Research Council (@StartingGrantERC), Junior project and Junior Star by Czech Science Foundation (2020), Walter N. Benjamin Programme by DFG (@WalterBenjaminProgramme), Innovation Research Incentives Scheme by Netherlands Organization of Scientific Research (NWO) (2021)... something from the US would be also good)

Although researcher independence seems to be important feature that is acknowledged by many actors in the research systems and that research funders try to stimulate, there hasn’t been much effort focused on quantifying this phenomenon. Some studies try to measure independence in a simple way, as a ratio of publications the given researcher produced without their supervisor divided by the total number of publications of the given researchers (e.g. @rojkoScientificPerformanceResearch2022). However, such indicator could be misleading in the world of team science, and also only focuses on collaborations, not tracking whether given researcher become independent in their thinking and research topics. [@vandenbesselaarMeasuringResearcherIndependence2019] therefore proposed more complex indicator for measuring researcher independence on the individual level, based on co-authorship network and topical overlaps. Their measure uses only information about publications’ characteristics (authorship and reference list) regardless of it’s impacts, unlike other alternative measures that focus solely on the impact of the independent work in terms of received citations (Dey, et al., 2021, @ALISNovelMetric).

Since the explicit goal of many funding instruments that are being used by research funders all over the world is to help early career researchers become independent, we would like to use the newly created indicators to assess the extent, to which such programmes influence researchers' independence.

@@

In the second step, we manually searched for PhD advisors of these 341 researchers in the databases of theses and dissertations that cover most universities in Czechia (https://theses.cz and https://dspace.cuni.cz/). We were able to match `r nrow(tar_read(ids_complete))` researchers with a PhD advisor (69,2 % of the Czech subsample). Further, we have not managed to find a suitable matched pair for another `r nrow(tar_read(ids_complete))-nrow(tar_read("matching") %>% filter(treatment == 1) %>% distinct(vedidk))` researchers in the matching procedure, reducing our final sample to `r nrow(tar_read("matching") %>% filter(treatment == 1) %>% distinct(vedidk))`.

@@

These researchers and their PhD advisors with their full publication history (limited to publications created at Czech institutions) created our final sample.

We also tested an assumption that had a hope of reducing time intensity of data collection by substitution it for algorithmic approach: using the "most frequent co-author of the first 5 publications" as a proxy to identify supervisors. Crosschecking this proxy with the actual data about supervisors we found manually, we found that this algorithm was able to identify supervisors correctly in about XXXXX % of cases (see the Attachment 1).

@@

However, we have made some changes to the indicator. After consulting with its authors, we have changed the aggregation formula to 
$$RII = ((1- I1) + I2 + I3 + I4) / 4$$ 
since the previous formula had a small error (it weighted I4 double,, i.e. 2*I4). In determining the proportion of a researcher's own research topics (I4), we opted for text-based topic modeling utilizing publication titles and abstracts, diverging from the originally suggested bibliographic coupling method. Initially, we conducted a random sampling of 1200 publications from a set of all publications produced by researchers affiliated with Czech institutions as cataloged in the RIV database, estimated number of topics using Griffiths2004 indicator and calculated LDA topic model on this set of publications. Then we  compiled all publications generated by the researchers within our sample and their respective supervisors. We then calculated the posterior distribution of topics for this compilation, employing a pre-trained topic model. This way we wanted to mitigate the endogeneity of the topic modelling which would be stronger if we calculated topic model directly on the publications of our sample researchers.

@@

We used the highest level of the OECD classification of disciplines, containing categories Natural Sciences; Engineering and Technology; Medical and Health Sciences; Agricultural Sciences; Social Sciences; Humanities. We chose this classification because it is widely used and also used in the dataset we worked with. For the matching we used the 42 sub-categories of this classification to achieve more granularity.

@@

We used Genderize.io API to estimate the probability of perceived gender based on the first names of researchers. Where available, information about the nationality of researchers was included in the API call to improve accuracy. 
We have also manually specified gender for 2 obrsevtation from the treatment group for which the gender was not calculated automatically.

@@

We analysed the data in the R version 4.2.3. using R studio and [this code](https://github.com/david-janku/juniors). For the construction of the independence indicator, we used network modelling (maybe more specific?) and text-based topic modelling (maybe more specific??).

For the construction of the control group, we implemented propensity matching using the package MatchIt (citation?) and matching exactly by the discipline (42 sub-categories of OECD classification) and treatment year, and approximately by number of publications in total, number of publications in WoS/Scopus, career age, interdisciplinary proportion, number of grants received prior to Junior Grant, career age when receiving the first grant and gender. All of these variables were calculated at the intervention year (i.e. year when the given researcher from experimental group received the Junior Grant). The distance was set to "mahalanobis". The matching was done in 1:3 ratio and with replacement. Manual search for supervisors of the matched control group observations was done by ranking the control group observation by distance from the treatment group observation and first searching for the supervisor of the closest control group observation, and if it was not possible to find it, then repeat the same for the second closest, etc... In the end each treatment observation was matched with exactly one control group observation which was closest but at the same time for which it was possible to identify their supervisor.

We have created 2 control groups: one using the matching above and one similar but specifically containing only researchers who received their first grant (other than Junior Grant) around the same career age as researchers in our intervention group. That allowed us to better track the impact of any grant funding vs specifically Junior Grants funding on researchers’ independence.

The difference between groups was tested using paired t-test and subsequent differences between discipline groups was tested using ANOVA post-hoc Tukey test.

@@

The independence indicator was not possible to calculate for `r na_counts` (`r na_percentage` % of observations). This is in `r na_sup` % caused by missing data about supervisors, and in `r na_pubs` % due to given researcher not having any publications in the given period (`r na_pubs_after` % of these cases are in the period after the intervention, which sometimes mean in the past few years).

@@

When Looking at disciplines, we see that there are slightly more missing data in Agricultural and veterinary sciences and the Humanities and the Arts.

@@

# Remove unnecessary columns and rename the others
df_summary <- df_summary %>%
  select(field, count_Complete, count_Missing, Proportion) %>%
  rename(
    Complete = count_Complete,
    Missing = count_Missing,
    Proportion_Missing = Proportion
  )

# Print the updated table
kable(df_summary, "html", digits = 3, caption = "Counts and Proportions of Missing Data by Field") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

@@

combined_table_before <- cbind(table_before_1, table_before_0[, -1], table_before_2[, -1])

>>>>

---
title: "Funding programmes for early career researchers do not improve their independence"
author:
- name: David Janků
  affiliation: '[1] Institute of Sociological Studies, Faculty of Social Sciences,
    Charles University, Czech Republic; [2] Centre for Science, Technology, and Society
    Studies, Institute of Philosophy, Czech Academy of Sciences, Czech Republic'
- name: Radim Hladík
  affiliation: '[1] Centre for Science, Technology, and Society Studies, Institute
    of Philosophy, Czech Academy of Sciences, Czech Republic'
output: bookdown::word_document2
always_allow_html: true
  # pdf_document: default
  # html_document:

@@

This work was supported by the Czech Science Foundation (project no. GJ20-01752Y, Funded and Unfunded Research in the Czech Republic: Scientometric Analysis and Topic Modeling)

@@

Name: David Janků 
Affiliation: 
[1] Institute of Sociological Studies, Faculty of Social Sciences, Charles University, Czech Republic
[2] Centre for Science, Technology, and Society Studies, Institute of Philosophy, Czech Academy of Sciences, Czech Republic  
Address: U Kříže 661/8, Praha 5 – Jinonice  
E-mail: david.janku@fsv.cuni.cz

@@

Independence is important quality constituting one factor of research excellence, often used in assessments for hiring, tenure and promotion in academia [@moherAssessingScientistsHiring2018]. From a bigger picture view, encouraging more independence might also lead to faster exploration of the cognitive map of science and thus lead to faster innovation. Many national research funders have created programmes specifically aimed at increasing independence of early career researchers, yet until recently, there was no quantifiable way to measure changes in independence. We employed recently proposed measure by @vandenbesselaarMeasuringResearcherIndependence2019 to assess whether one such a funding programme by Czech Science Foundation is effective at stimulating more independence. We found that funding for early career researchers is not effective at stimulating their independence, regardless of the source of funding. The strongest effect on independence growth is time, explaining around 60% of the variance in independence.  Additionally, we found differences in levels of independence across disciplines that were stable across time and treatment groups, with Social Sciences and Humanities showing highest independence, while  Agricultural and Veterinary sciences and Medical and Health Sciences showing lowest independence. However, these differences reflect only changes in collaboration patterns, while there were no changes in cognitive patterns across disciplines and only modest changes across time (from 40 % to 46 % of topics not shared by their PhD advisor).

## Keywords
research independence; early career researchers; grant funding; mentoring; counter-factual analysis; research evaluation

# Introduction

To what extent is scientific progress conserved and impeded by novices mirroring seniors’ research agendas? Such a phenomenon is likely to occur because following tradition is usually a more beneficial and safer option for novices. Indeed, empirical data suggest that this strategy is chosen by a majority of researchers [@foster2015], and only a small fraction dares to deviate and explore the cognitive map of science more rapidly. For early career researchers specifically, partnering with already well-established seniors provides a large career boost [@li2019] and being thematically closer to their PhD as well as their post-doc advisor increases the probability of continuing in academia and having more mentees later on [@lienard2018]. However, this effect might stifle scientific progress as it leads to a slower exploration of its cognitive map [@rzhetsky2015]. Moreover, this mirroring further strengthens the existing positions of senior faculty, making little space for disrupting the status quo and creating new perspectives.
This was illustrated by @azoulay2010, who investigated the effects of unexpected deaths of prominent scientists in their research fields and their collaborators' careers. They found a decrease in the productivity of junior collaborators and an increase in published work by young newcomers in the field rewarded by an increased citation rate. That suggests the sudden departures of prominent scientists gave room for more disruptive changes in the field and for bringing new ideas and perspectives while harming the status quo holders who worked with the prominent scientists on their research agendas - a “Goliath effect,” as coined by @wang2021, p.42.

This mirroring effect is also closely connected to the concept of academic inbreeding [@machacek2022], which focuses more on academic mobility and conservation of institutional habitus rather than specifically on the inheritance of research agendas. However, inbreeding is also often thought to increase dependence and decrease exploration of new ideas.

As younger researchers are already more likely to come up with new topics and ideas [@packalen2019], it would be fitting to empower them and stimulate their independence so that they can “challenge the existing common views and open up new lines of research” [@vandenbesselaarMeasuringResearcherIndependence2019, p.2] and accelerate scientific progress rather than recreate the existing status quo. An independent investigator is then defined by the US National Academies of Science [@nationalresearchcouncilus2005] as “one who enjoys the independence of thought—the freedom to define the problem of interest and to choose or develop the best strategies and approaches to addressing that problem–especially during the earlier career phase” [@vandenbesselaarMeasuringResearcherIndependence2019, p.4]. Such a definition does not necessarily put independence in relation to other actors (e.g. one’s PhD advisor), but only in relation to one’s own freedom and resources. However other institutions, such as European Research Council via their Starting Grants scheme [@neufeldPeerReviewbasedSelection2013] think of independence in a relational manner, in terms of having one (or a few) publications without the former PhD supervisor being co-author; being PI on a grant; or being lab director or research leader.

Developing independence is also one of the things that the system of science expects from its members - it is one of the criteria often used for assessing researchers for hiring, tenure and promotion [@vanarensbergenDifferentViewsScholarly2014; @moherAssessingScientistsHiring2018], constituting one factor of research excellence. Even from the perspective of scholarly output, protégés were found to “do their best work when they break from their mentor’s research topics and coauthor no more than a small portion of their overall research with their mentors” [@ma2020]. @lienard2018 found that making a larger thematic step from the PhD advisor and finding a post-doc advisor with less similar interests was also related to increased chances of continuing in academia and having more mentees later on (although under the condition that one remains thematically close to the post-doc advisor). There is also some direct preliminary evidence that higher independence modestly correlates with the number of citations acquired [@rojkoScientificPerformanceResearch2022], but others argue that even the research systems with very low independence can create a lot of scientific impact, and thus the relationships between independence and impact might not be as clear [@glaserIndependenceResearchReview2022].

However, developing this independence does not go without saying. There are few known factors influencing the level of research independence. One of them is study discipline. Based on some early empirical studies [@rojkoScientificPerformanceResearch2022], it seems that researchers in social sciences and humanities experience significantly higher levels of independence from their PhD advisor than researchers in other disciplines.

Interestingly, there is also some evidence that the propensity to cognitive independence might be inherited and passed down from generation to generation in mentor-mentee relationships [@yoshioka-kobayashi2021] - perhaps it is not only research agendas that students inherit from their advisors but also the tendency to deviate from those research agendas.

In recent decades, there has been a worry that researchers’ independence is decreasing, illustrated by an increase in the age at which researchers received their first grant [@powell2016; @christian2021; and @nationalresearchcouncilus2005; @danielsGenerationRiskYoung2015 for US biomedical researchers specifically], as well as direct measures of new researchers becoming more dependent on their PhD advisors across almost all disciplines in Slovenia [@rojkoScientificPerformanceResearch2022, table 6 on p.9]. These trends have motivated some to create new proposals for funding agencies and other research funders in order to support early career researchers [@dewinde2021]

Perhaps to tackle this trend, many grant agencies decided to support early career researcher by funding. Some have decided to address this trend by giving advantage to first-awardees and early career researchers in their usual distribution process, while others have created specific programmes dedicated to funding young researchers, typically up to 8 years after completing their PhD. Such programmes have often explicitly stated goals of helping researchers become more independent, set up their own research groups and open up new lines of research. Examples of such programmes could be Starting Grants by European Research Council [@StartingGrantERC], Junior project and Junior Star by @czechsciencefoundation2019, Walter N. Benjamin Programme by DFG [@WalterBenjaminProgramme], Innovation Research Incentives Scheme by Netherlands Organization of Scientific Research [@nwo2023].

Grant funding mechanisms have previously been shown to stimulate productivity in terms of the number of publications and impact in terms of citations [@gallo2018; for the Czech environment specifically see @bajgar2021; notable exception is @vandenbesselaar2015, who specifically looked at funding programmes for early career researchers and found small or no effect], and even scientific novelty [@wang2018, or, as some authors argue, rather interdisciplinarity: @fontana2020]. In terms of career progress, obtaining grant funding was connected to significantly higher chances of career progression and obtaining a professorship [@bloch2014, @vandenbesselaar2015].

Funding is thus thought to be one of the mechanisms that could stimulate researchers' independence. However, there hasn’t been much effort focused on quantifying this phenomenon. Some studies try to measure independence in a simple way, as a ratio of publications the given researcher produced without their supervisor divided by the total number of publications of the given researchers [e.g. @rojkoScientificPerformanceResearch2022]. However, such an indicator could be misleading in the world of team science, and also only focuses on collaborations, not tracking whether a given researcher becomes independent in their thinking and research topics. @vandenbesselaarMeasuringResearcherIndependence2019 therefore proposed a more complex indicator for measuring researcher independence on the individual level, based on co-authorship network and topical overlaps. Their measure uses only information about publications’ characteristics (authorship and reference list) regardless of their impacts, unlike other alternative measures that focus solely on the impact of the independent work in terms of received citations [@dey2023].

Understanding the impact of early-career funding on research independence is crucial for shaping future funding policies and, ultimately, the trajectory of scientific innovation. This study serves as a pivotal step in quantifying that impact, providing actionable insights for funding agencies and academic institutions alike.

@@

In the second step, we manually searched for PhD advisors of these 341 researchers in the databases of theses and dissertations that cover most universities in Czechia (https://theses.cz and https://dspace.cuni.cz/). We were able to match `r nrow(tar_read(ids_complete) %>% filter(!is.na(sup_name)))` researchers with a PhD advisor (82 % of the Czech subsample) and find suitable matched pair from control group for each of them. 
<!-- Further, we have not managed to find a suitable matched pair for another `r nrow(tar_read(ids_complete))-nrow(tar_read("matching") %>% filter(treatment == 1) %>% distinct(vedidk))` researchers in the matching procedure, reducing our final sample to `r nrow(tar_read("matching") %>% filter(treatment == 1) %>% distinct(vedidk))`.  -->

@@

These researchers and their PhD advisors with their full publication history (limited to publications created at Czech institutions) constituted our final sample.

We also tested an assumption that had a hope of reducing time intensity of data collection by substituting it with an algorithmic approach: using the "most frequent co-author of the first 5 publications" as a proxy to identify supervisors. Crosschecking this proxy with the actual data about supervisors we found manually showed that this algorithm was able to identify supervisors correctly in only about 56 % of cases (see the Attachments).

@@

However, we have made some changes to the indicator. After consulting with its authors, we have changed the aggregation formula to 
$$RII = ((1- I1) + I2 + I3 + I4) / 4$$ 
since the previous formula had a small error (it weighted I4 double,, i.e. 2*I4). In determining the proportion of a researcher's own research topics (I4), we opted for text-based topic modeling utilizing publication titles and abstracts, diverging from the originally suggested bibliographic coupling method. Initially, we conducted a random sampling of 3000 publications from a set of all publications produced by researchers affiliated with Czech institutions as cataloged in the RIV database, estimated number of topics using Griffiths2004 indicator, and calculated LDA topic model on this set of publications. Then we  compiled all publications generated by the researchers within our sample and their respective supervisors. We then calculated the posterior distribution of topics for this compilation, employing a pre-trained topic model. This way we wanted to mitigate the endogeneity of the topic modelling which would be stronger if we calculated topic model directly on the publications of our sample researchers.

To calculate this indicator, we required at least 3 publications. This led to an increased number of missing observations (see section Descriptives - Missing data).

@@

We used the highest level of the OECD classification of disciplines: containing categories Natural Sciences; Engineering and Technology; Medical and Health Sciences; Agricultural Sciences; Social Sciences; Humanities. We chose this classification because it is widely used and also present in the dataset we worked with. For the matching we used the 42 sub-categories of this classification to achieve more granularity.

@@

We used Genderize.io API to estimate the probability of perceived gender based on the first names of researchers. We had also manually specify gender for 2 observations from the treatment group for which the gender was not calculated automatically.

@@

We analysed the data in the R version 4.2.3. using R studio and [this code](https://github.com/david-janku/juniors). For the construction of the independence indicator, we used network modelling and text-based topic modelling.

For the construction of the control group, we implemented propensity matching using the package MatchIt [@ho2011] and matching exactly by the discipline (42 sub-categories of OECD classification) and treatment year, and approximately by number of publications in total, number of publications in WoS/Scopus, career age, interdisciplinary proportion, number of grants received prior to Junior Grant, career age when receiving the first grant, and gender. All of these variables were calculated in the intervention year (i.e. year when the given researcher from experimental group received the Junior Grant). The distance was set to "mahalanobis". The matching was done in 1:1 ratio and with replacement.

Using the matching algorithm, we have created 2 control groups: "Unfunded" group did not receive any grants prior to and 3 years after the treatment year. "Funded" group was matched based on career year at which they received their first grant, ensuring that while treatment group researchers received Junior Grant, "funded" researchers received a different grant at the same time. That allowed us to better track the impact of any grant funding vs specifically Junior Grants funding on researchers’ independence.

The difference between treatment groups was tested using repeated measures ANOVA with a special error term taking into account paired nature of the data. The ANOVA model was specified as RII ~ treatment * independence_timing, incorporating both the main effects and the interaction between treatment and independence timing. A significant aspect of our ANOVA approach was the inclusion of an error term structured as Error(subclass/independence_timing). This error structure accounts for the nested or hierarchical nature of our data, where measurements are nested within subclasses and vary across different independence timings. By doing so, the model adjusts for potential within-subclass correlations and differences across independence timings.

To ascertain the significance of academic field as a predictor, we constructed two additional GLS models. The first model included treatment, independence timing, and their interaction as predictors (treatment + independence_timing + treatment:independence_timing), while the second model added the main effect of the academic field (+ field). Both models employed a compound symmetry correlation structure (corCompSymm) based on subclass and independence timing, suitable for data with equal correlation between observations. By comparing these models, we aimed to evaluate the impact of including the academic field as a predictor on the model's explanatory power, thereby testing its statistical significance in explaining variations in RII.

@@

The independence indicator was not possible to calculate for `r na_counts` (`r na_percentage` % of observations). This is in `r na_sup` % caused by missing data about supervisors, and in `r na_pubs` % due to absence of publications by the researcher during the specified period (`r na_pubs_after` % of these cases occurred post-intervention, often in the past few years).

@@

From a perspective of disciplines, we see that there are slightly more missing data in Agricultural and veterinary sciences and the Humanities and the Arts.

@@

# Remove unnecessary columns and rename the others
df_summary <- df_summary %>%
  select(field, count_Complete, count_Missing, Proportion) %>%
  rename(
    Field = field,
    Complete = count_Complete,
    Missing = count_Missing,
    Proportion_Missing = Proportion
  )

# Print the updated table
kable(df_summary, digits = 3, caption = "Counts and Proportions of Missing Data by Field") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

@@

combined_table_before <- cbind(table_before_1, table_before_0, table_before_2)

kable(combined_table_before, 
      digits = 3, align = "c", booktabs = TRUE, caption = "Descriptives of All Groups Before Intervention") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  add_header_above(c(" " = 1, "Treatment Group" = 5, "Unfunded Control Group" = 5, "Funded Control Group" = 5))
